---
title: "preliminary_mlm.Rmd"
author: "Mingxuan Liu7"
date: "11/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Item level IV1: visual cues (2 levels: healthy food and junk food)
Item level IV2: verbal cues (2 levels: encouragement frame and discouragement frame)
Equation for the appetitive response (DV, indexed by oo); 
$$
\begin{aligned}
  \text{oo}_{i(j,k)} & = \gamma_{00} \\ 
                        & + \gamma_{10} \text{visual}_{k} + \gamma_{20} \text{verbal}_{k} + \gamma_{30} \text{time}_{k}\\
                        & + \gamma_{40} \text{visual}_{k} \times \text{time}_{k} + \gamma_{50} \text{verbal}_{k} \times \text{time}_{k} + \gamma_{60} \text{visual}_{k} \times \text{verbal}_{k}+ \gamma_{70} \text{visual}_{k} \times \text{verbal}_{k}\times \text{time}_{k}\\
                        & + u_{0j} + u_{1j} \text{visual}_{k} + u_{2j} \text{verbal}_{k} + u_{3j} \text{time}_{k} +  u_{4j} \text{visual}_{k} \times \text{verbal}_{k}+ u_{5j} \text{visual}_{k} \times \text{verbal}_{k}\times \text{time}_{k} \\
                        & + e_{ijk}
\end{aligned}
$$



```{r load-pkg, message=FALSE}
# To install a package, run the following ONCE (and only once on your computer)
# install.packages("psych")  
library(here)  # makes reading data more consistent
library(tidyverse)  # for data manipulation and plotting
library(haven)  # for importing SPSS/SAS/Stata data
library(lme4)  # for multilevel analysis
library(lmerTest)  # for testing coefficients
library(MuMIn)  # for R^2
library(sjPlot)  # for plotting effects
library(emmeans)  # for marginal means
library(modelsummary)  # for making tables
library(dplyr)
library(psych)
theme_set(theme_bw())  # Theme; just my personal preference
```

、

```{r import data}
oo <- read_sav(here("data_files", "oo_survey.sav"))
oo
colnames(oo)

corr <- read_sav(here("data_files", "corr_survey.sav"))
corr
colnames(corr)
```
8 items:
h_do_drink
h_do_snack
h_dont_drink
h_dont_snack
j_do_drink
j_do_snack
j_dont_drink
j_dont_snack
time1*6
time2*12
time3*6

```{r show-labels}
# Show the variable labels
cbind(
    map(oo, attr, "label")
)
```

```{r wide to long, echo=FALSE}

oo_long <- oo %>%
  select(2, 9:20, 33:44, 57:68, 81:92, 105:116, 129:140, 153:164, 177:188, 195:232)%>%
  pivot_longer(
    c(h_do_drink_t1:h_do_drink_t12, j_do_snack_t1:j_do_snack_t12,h_dont_snack_t1:h_dont_snack_t12,j_dont_drink_t1:j_dont_drink_t12, h_do_snack_t1:h_do_snack_t12, j_do_drink_t1:j_do_drink_t12, h_dont_drink_t1:h_dont_drink_t12, j_dont_snack_t1:j_dont_snack_t12),  # variables that are repeated measures
    # Convert 8 columns to 3: 2 columns each for anti/read (.value), and
    # one column for time
    names_to = c("viscue", "verbcue", "type", "time"),
    # Extract the names "anti"/"read" from the names of the variables for the
    # value columns, and then the number to the "time" column
   names_pattern = "(h|j)_(do|dont)_(drink|snack)_t(.*)",
    # Convert the "time" column to integers
    names_transform = list(time = as.integer)
  )%>%
unite(item, viscue, verbcue, type, remove = FALSE)
oo_long 

corr_long <- corr %>%
  select(2, 9:20, 33:44, 57:68, 81:92, 105:116, 129:140, 153:164, 177:188, 195:232)%>%
  pivot_longer(
    c(h_do_drink_t1:h_do_drink_t12, j_do_snack_t1:j_do_snack_t12,h_dont_snack_t1:h_dont_snack_t12,j_dont_drink_t1:j_dont_drink_t12, h_do_snack_t1:h_do_snack_t12, j_do_drink_t1:j_do_drink_t12, h_dont_drink_t1:h_dont_drink_t12, j_dont_snack_t1:j_dont_snack_t12),  # variables that are repeated measures
    # Convert 8 columns to 3: 2 columns each for anti/read (.value), and
    # one column for time
    names_to = c("viscue", "verbcue", "type", "time"),
    # Extract the names "anti"/"read" from the names of the variables for the
    # value columns, and then the number to the "time" column
   names_pattern = "(h|j)_(do|dont)_(drink|snack)_t(.*)",
    # Convert the "time" column to integers
    names_transform = list(time = as.integer)
  )%>%
unite(item, viscue, verbcue, type, remove = FALSE)
corr_long 

write.csv(oo_long, "oo_long.csv")
oo_long <- read.csv(here("data_files", "oo_long_withimage.csv"))

write.csv(corr_long, "corr_long.csv")
corr_long <- read.csv(here("data_files", "corr_long.csv"))

oo_corr <- read.csv(here("data_files", "oo_corr_long.csv"))
```

```{r models, echo=FALSE}
m1 <- lmer(corr ~  (1 | id) + (1 | image),
           data = oo_corr)
ranova(m1) #all sig.
m2 <- lmer(corr ~ verbcue+time + (1| id)+ (1 | image),
           data = oo_corr)
ranova(m2)
summary(m2)
m3 <- lmer(value ~ viscue * verbcue * time + (verbcue | id) + (1 | image),
           data = oo_long)
ranova(m3)#all sig.
m4 <- lmer(value ~ viscue * verbcue * time +  (viscue | id) + (1 | image),
           data = oo_long)
ranova(m4)#all sig.; failed to converge

m5 <- lmer(value ~ viscue * verbcue * time +  (time | id) + (1 | image),
           data = oo_long)
ranova(m5)#all sig.; failed to converge

m6 <- lmer(value ~ viscue * verbcue * time + (viscue | id) + (verbcue | id) + (time | id) + (1 | image),   data = oo_long)
ranova(m6)


summary(m5)
msummary(list(
  "random intercept" = m1,
  "full model" = m5
))
```

```{r Healthy subset, echo=FALSE}
h <- oo_corr[which(oo_corr$viscue=='h'), ]
h

h$time2 <- h$time * h$time
h$time3 <- h$time2 * h$time

h1 <- lmer(oo ~  time + (1 | id) + (1 | image),
           data = h)
summary(h1) #all sig.

h2 <- lmer(corr ~ time + (1| id)+ (1 | image),
           data = h)
summary(h2)


h_time <- aggregate(h$value~h$time, FUN=mean)
names(h_time)[names(h_time) == "h$time"] <- "time"
names(h_time)[names(h_time) == "h$value"] <- "value"

p1 <- ggplot(data=h_time, aes(x=time, y=value)) +
  geom_line()+
  geom_point() 
p1
p1 + scale_x_continuous(breaks = seq(1, 12, by = 1))
```

```{r Junk subset, echo=FALSE}
j <- oo_corr[which(oo_corr$viscue=='j'), ]
j

j$time2 <- j$time * j$time
j$time3 <- j$time2 * j$time

j1 <- lmer(oo ~  time + (1 | id) + (1 | image),
           data = j)
summary(j1) #all sig.

j2 <- lmer(corr ~  time + (1 | id) + (1 | image),
           data = j)
summary(j2)
```
```{r Do subset, echo=FALSE}
do <- oo_corr[which(oo_corr$verbcue == 'do'), ]
do

do$time2 <- do$time * do$time
do$time3 <- do$time2 * do$time

do1 <- lmer(oo ~  time + (1 | id) + (1 | image),
           data = do)
summary(do1) #all sig.

do2 <- lmer(corr ~  time + (1 | id) + (1 | image),
           data = do)
summary(do2)
```
```{r Dont subset, echo=FALSE}
dont <- oo_corr[which(oo_corr$verbcue == 'dont'), ]
dont

dont$time2 <- dont$time * dont$time
dont$time3 <- dont$time2 * dont$time

dont1 <- lmer(oo ~  time + (1 | id) + (1 | image),
           data = dont)
summary(dont1) #all sig.

dont2 <- lmer(corr ~  time + (1 | id) + (1 | image),
           data = dont)
summary(dont2)
```


```{r Healthy X Do subset, echo=FALSE}
h_do <- oo_long[which(oo_long$viscue=='h'
& oo_long$verbcue == 'do'), ]
h_do

h_do$time2 <- h_do$time * h_do$time
h_do$time3 <- h_do$time2 * h_do$time

hdo1 <- lmer(value ~  (1 | id) + (1 | image),
           data = h_do)
ranova(hdo1) #all sig.

hdo2 <- lmer(value ~ time + (1| id),
           data = h_do)
summary(hdo2)
```

```{r Junk X Dont subset, echo=FALSE}
j_dont <- oo_long[which(oo_long$viscue=='j'
& oo_long$verbcue == 'dont'), ]
j_dont

j_dont$time2 <- j_dont$time * j_dont$time
j_dont$time3 <- j_dont$time2 * j_dont$time

jdont1 <- lmer(value ~  (1 | id) + (1 | image),
           data = j_dont)
ranova(jdont1) #all sig.

jdont2 <- lmer(value ~ time + time2+ time3+
                (1| id)+ (1 | image),
           data = j_dont)
summary(jdont2)
```

```{r Healthy X Dont subset, echo=FALSE}
h_dont <- oo_long[which(oo_long$viscue=='h'
& oo_long$verbcue == 'dont'), ]
h_dont

h_dont$time2 <- h_dont$time * h_dont$time
h_dont$time3 <- h_dont$time2 * h_dont$time

hdont1 <- lmer(value ~  (1 | id) + (1 | image),
           data = h_dont)
ranova(hdont1) #all sig.

hdont2 <- lmer(value ~ time + time2+
                (1| id)+ (1 | image),
           data = h_dont)
summary(hdont2)

####how to interpret the sig. after higher level time was added
####how to get effect size
```

```{r Junk X Do subset, echo=FALSE}
j_do <- oo_long[which(oo_long$viscue=='j'
& oo_long$verbcue == 'do'), ]
j_do

j_do$time2 <- j_do$time * j_do$time
j_do$time3 <- j_do$time2 * j_do$time

jdo1 <- lmer(value ~  (1 | id) + (1 | image),
           data = j_do)
ranova(jdo1) #all sig.

jdo2 <- lmer(value ~ time + time2 + time3 + (1| id),
           data = j_do)
summary(jdo2)
```

```{r plot, echo=FALSE}
p1 <- ggplot(j_dont, aes(x = time, y = value)) + 
  geom_point() + 
  geom_line(aes(group = id)) +  # add lines to connect the data for each person
  # add a mean trajectory
  stat_summary(fun = "mean", col = "red", size = 1, geom = "line")
p1
```
The model showed a main effect of time on participants' appetitive activation (indexed by oo value). Significant random slopes of visual cues, verbal cues, and time across persons indicated that the effect of visual cues, verbal cues, and time varied across participants. 


####double check mean of each category
```{r check mean, echo=FALSE}
mean(data.matrix(oo_long[oo_long$viscue == "h","value"]))
mean(data.matrix(oo_long[oo_long$viscue == "j","value"]))
mean(data.matrix(oo_long[oo_long$verbcue == "do","value"]))
mean(data.matrix(oo_long[oo_long$verbcue == "dont","value"]))
mean(data.matrix(oo_long[oo_long$time == "1","value"]))
mean(data.matrix(oo_long[oo_long$time == "2","value"]))
mean(data.matrix(oo_long[oo_long$time == "3","value"]))
mean(data.matrix(oo_long[oo_long$time == "4","value"]))
mean(data.matrix(oo_long[oo_long$time == "5","value"]))
mean(data.matrix(oo_long[oo_long$time == "6","value"]))
mean(data.matrix(oo_long[oo_long$time == "7","value"]))
mean(data.matrix(oo_long[oo_long$time == "8","value"]))
mean(data.matrix(oo_long[oo_long$time == "9","value"]))
mean(data.matrix(oo_long[oo_long$time == "10","value"]))
mean(data.matrix(oo_long[oo_long$time == "11","value"]))
mean(data.matrix(oo_long[oo_long$time == "12","value"]))
```